---
title: "Final_Project"
author: "Songyi Ahn, Nakyung Lee, Nishat Parvez"
format:
  html:
   highlight-style: tango
   code-fold: true
   toc: true
   toc-depth: 5
   toc-location: left
   self-contained: true
execute:
  echo: true
  eval: true
  warning: false
  message: false
bibliography: finalprojectreferences.bib 
editor: visual
---

## Abstract

**(please write abstract)**

## Introduction

The increasing adoption of telehealth services has transformed healthcare delivery by offering more accessible and convenient solutions for diverse health needs [@weinstein2014]. However, ensuring equitable access to these services requires a deeper understanding of the characteristics that influence their utilization. This study focuses on identifying and analyzing demographic, socio-economic, and behavioral factors that differentiate telehealth users from non-users with an exploratory object regarding this matter. By examining these factors, the study seeks to provide rudimentary insights into how telehealth can better serve diverse populations.

This research addresses three key areas by leveraging the Health Information National Trends Survey data from the National Cancer Institute [-@hints]. First, it examines demographic and socio-economic factors to uncover patterns in telehealth usage and potential disparities in digital health engagement because individual characteristics could serve as barriers or facilitators for telehealth access [@rani2019]. Second, it explores social media usage patterns and engagement levels with health information, comparing telehealth users and non-users. Given the growing role of social media as a source of health information [@zhao2017], social media usage and activities regarding health information could be some meaningful factors in understanding people's willingness to adopt digital healthcare services. Finally, the study investigates cancer risk perception among telehealth users and non-users to analyze how health awareness is related to telehealth usage. Since telehealth can be an effective tool for health care or education [@garber2022], grasping how people perceive their cancer risk could help tailor telehealth services to meet the needs of different groups of people. By integrating these dimensions, the research could contribute to a deeper understanding of telehealth adoption and their role in enhancing digital health equity for public health.

## Research Questions

-   *RQ1: How do demographic and socio-economic factors differ between telehealth users and non-users?*

-   *RQ2: How do social media usage frequency and health information engagement levels differ between telehealth users and non-users?*

-   *RQ3: How do cancer-related risk perceptions differ between telehealth users and non-users?*

## Methods

**(please explain what data we used, when it was collected, by whom etc...).**

The research question, "How do demographic and socio-economic factors differ between telehealth users and non-users?" aligns closely with several survey questions from the HINTS data, focusing on telehealth usage, demographic factors such as age, gender, race and ethnicity, education, and household income, as well as socio-economic factors like employment, marital status, and financial comfort. These variables provide a comprehensive framework for analyzing patterns and disparities in telehealth adoption. The data preparation process involved selecting and recoding these variables to enhance interpretability, with missing or ambiguous responses replaced with NA. Ethnicity was harmonized by combining Hispanic and racial data for a unified classification, while educational attainment and income levels were grouped into broader categories. Telehealth usage was simplified into binary categories (e.g., "Yes" or "No"), and unnecessary columns were dropped, with rows containing missing data excluded to ensure completeness. The dataset was then transformed into a long format for targeted filtering and analysis before reverting to its original structure for interpretative consistency. This systematic approach ensures a clean, reliable dataset to examine demographic and socio-economic differences in telehealth usage, contributing to a deeper understanding of accessibility and equity in digital healthcare solutions.

**(Please add your own measures here Nakyung!)**

Telehealth usage was based on responses to a survey question asking whether the participant had received care from a doctor or health professional through telehealth in the past year, with options of "yes" or "no."

Cancer risk perception was measured by averaging responses to a questionnaire that asked about the participant's interest in receiving cancer screenings in the coming year, their fear of developing cancer, how susceptible they felt to cancer, their belief in the prevalence of cancer-causing factors, and their opinion on the effectiveness of cancer prevention activities. Responses to these questions were converted into a 5-point Likert scale. Based on the average score, individuals were classified into three groups: those in the lowest third of scores were categorized as "low level," those in the middle third as "medium level," and the remaining individuals as "high level."

For this project, program R created by @r was used for data analysis. Then R packages such as tidyverse [@tidyverse], here [@here], and rio [@rio] were applied.

```{r}
# loading pacakages
#| echo: false
#| include: false

library(tidyverse)
library(here)
library(rio)
library(knitr)
library(tinytex)
```

```{r}
# Load the file
file_path <- here("data", "hints6_public.rda")
load(file_path)

# Change the file name to 'hints'
hints <- public
```

## Result

### Demographic Factors and Telehealth

Telehealth usage was positively associated with younger age, higher education levels, and greater income, while disparities were observed among ethnic groups, with lower usage among "Black" and "Hispanic" participants.

#### RQ1_1. Data Cleaning

```{r}
#| echo: false

# Select relevant columns for RQ1
health_data <- hints %>%
  select(AgeGrpB, BirthGender, RACE_CAT2, HISP_CAT, Education, IncomeRanges, ReceiveTelehealthCare)

# Recode, rename variables, and handle missing data
health_data_clean <- health_data %>%
  mutate(
    # Clean the truncated value in the Education column
    Education = str_replace(Education, 
                             "Post high school training other than college.*", 
                             "Post high school training other than college (vocational)"),
    
    # Recode Telehealth
    Telehealth = recode(
      ReceiveTelehealthCare,
      "No telehealth visits in the past 12 months" = "No",
      "Yes, by video" = "Yes",
      "Yes, by phone call (voice only with no video)" = "Yes",
      "Yes, some by video and some by phone call" = "Yes",
      "Missing data (Web partial - Question Never Seen)" = NA_character_,
      "Missing data (Not Ascertained)" = NA_character_
    ),
    
    # Recode Age
    Age = recode(
      AgeGrpB,
      "1" = "18-24",
      "2" = "25-34",
      "3" = "35-44",
      "4" = "45-54",
      "5" = "55-64",
      "6" = "65-74",
      "7" = "75+",
      "Missing Data (Not Ascertained)" = NA_character_, 
      "Unreadable or Nonconforming Numeric Response" = NA_character_
    ),
    
    # Recode Gender
    Gender = recode(
      BirthGender,
      "1" = "Male",
      "2" = "Female",
      "Missing data (Not Ascertained)" = NA_character_
    ),
    
    # Recode Education
    Education = recode(
      Education,
      "Less than 8 years" = "Low Education",
      "8 through 11 years" = "Low Education",      
      "12 years or completed high school" = "Medium Education",
      "Post high school training other than college (vocational)" = "Medium Education",
      "Some college" = "High Education",
      "College graduate" = "High Education",         
      "Postgraduate" = "High Education",    
      "Missing data (Not Ascertained)" = NA_character_,
      "Missing data (Web partial - Question Never Seen)" = NA_character_
    ),
    
    # Recode Income
    Income = recode(
      IncomeRanges,
      "$0 to $9,999" = "<$10K",
      "$10,000 to $14,999" = "$10K-$14K",
      "$15,000 to $19,999" = "$15K-$19K",
      "$20,000 to $34,999" = "$20K-$34K",
      "$35,000 to $49,999" = "$35K-$49K",
      "$50,000 to $74,999" = "$50K-$74K",
      "$75,000 to $99,999" = "$75K-$99K",
      "$100,000 and $199,999" = "$100K-$199K",
      "$200,000 or more" = "$200K",
      "Missing data (Not Ascertained)" = NA_character_,
      "Missing data (Web partial - Question Never Seen)" = NA_character_
    ),
    
    # Recode Ethnicity
    Ethnicity = recode(
      HISP_CAT,
      "Mexican only" = "Hispanic",
      "Cuban only" = "Hispanic",
      "Puerto Rican only" = "Hispanic",
      "Other Hispanic only" = "Other Hispanic or Latino/a",
      "Not Hispanic only" = NA_character_,
      "Multiple Hispanic ethnicities selected" = NA_character_,
      "Missing data (Not Ascertained)" = NA_character_,
      "Missing data (Web partial - Question Never Seen)" = NA_character_
    ) %>% 
      coalesce(recode(
        RACE_CAT2,
        "White only" = "White",
        "Black only" = "Black",
        "Filipino only" = "Asian or Pacific Islander",
        "Chinese only" = "Asian or Pacific Islander",
        "Asian Indian only" = "Asian or Pacific Islander",
        "Other Asian only" = "Asian or Pacific Islander",
        "Asian only" = "Asian or Pacific Islander",
        "Vietnamese only" = "Asian or Pacific Islander",
        "Other Pacific Islander only" = "Asian or Pacific Islander",
        "American Indian or Alaska Native only" = "Native American",
        "Missing data (Not Ascertained)" = NA_character_,
        "Missing data (Web partial - Question Never Seen)" = NA_character_,
        "Multiple races selected" = NA_character_
      ))
  ) %>%
  # Drop unnecessary columns
  select(-AgeGrpB, -RACE_CAT2, -HISP_CAT, -BirthGender, -IncomeRanges, -ReceiveTelehealthCare) %>%
  na.omit()

# Transform data using pivot_longer and pivot_wider
# Pivot longer
health_data_long <- health_data_clean %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  filter(!is.na(Value))

# Pivot wider back to the original structure
health_data_wide <- health_data_long %>%
  pivot_wider(
    names_from = Variable,
    values_from = Value
  )

# View the cleaned dataset
head(health_data_clean)


```

#### RQ1_2: Data Visualization

```{r}
#| echo: false

# List of demographic variables to visualize
demographic_vars <- c("Age", "Gender", "Education", "Income", "Ethnicity")

# Function to create and display ggplot for each demographic variable
create_telehealth_plot <- function(data, var) {
  # Remove missing data for the current variable
  data <- data %>%
    filter(!is.na(!!sym(var)), !is.na(Telehealth)) %>%
    group_by(!!sym(var), Telehealth) %>%
    summarise(count = n(), .groups = "drop") %>%
    mutate(percentage = count / sum(count) * 100) # Calculate percentages
  
  ggplot(data, aes_string(x = var, y = "count", fill = "Telehealth")) +
    geom_bar(stat = "identity", position = "dodge", color = "black", width = 0.7) + # Bar chart
    geom_text(
      aes(label = paste0(round(percentage, 1), "%")), # Display percentages on bars
      position = position_dodge(width = 0.7), 
      vjust = -0.3, size = 3.5 # Adjust position and size of text
    ) +
    labs(
      title = paste("Telehealth Usage by", var),
      x = var,
      y = "Number of Responses",
      fill = "Telehealth Usage"
    ) +
    scale_fill_brewer(palette = "Set2") + # Use a colorblind-friendly palette
    theme_minimal(base_size = 14) + # Use a clean theme with larger base font size
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels for readability
      axis.title = element_text(face = "bold"), # Bold axis titles
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16), 
    # Center and bold title
      legend.position = "top", # Place legend at the top
      legend.title = element_text(face = "bold") # Bold legend title
    )}
# Loop through each demographic variable and create a plot
plots <- lapply(demographic_vars, function(var) {
  create_telehealth_plot(health_data_clean, var)})

# Display the plots
for (plot in plots) {
  print(plot)
}
```

#### RQ1_3: Summary Statistics

```{r}
#| echo: false

# Grouped summary for demographics and telehealth usage
summary_table_demographics <- health_data_clean %>%
  group_by(Telehealth, Age, Gender, Education, Income, Ethnicity) %>%
  summarize(count = n(), .groups = "drop") %>%
  mutate(percentage = round(count / sum(count) * 100, 2))

# Display the demographic summary table
kable(summary_table_demographics, caption = "Summary of Telehealth Usage by Demographics")

# Summary statistics for Telehealth usage
summary_table_telehealth_stats <- health_data_clean %>%
  group_by(Telehealth) %>%
  summarize(
    Age_mean = round(mean(as.numeric(Age), na.rm = TRUE), 2),
    Age_median = median(as.numeric(Age), na.rm = TRUE),
    Age_sd = round(sd(as.numeric(Age), na.rm = TRUE), 2),
    
    Education_mean = round(mean(as.numeric(Education), na.rm = TRUE), 2),
    Education_median = median(as.numeric(Education), na.rm = TRUE),
    Education_sd = round(sd(as.numeric(Education), na.rm = TRUE), 2),
    
    Income_mean = round(mean(as.numeric(as.factor(Income)), na.rm = TRUE), 2),
    Income_median = median(as.numeric(as.factor(Income)), na.rm = TRUE),
    Income_sd = round(sd(as.numeric(as.factor(Income)), na.rm = TRUE), 2)
  )

# Display the summary statistics table
kable(summary_table_telehealth_stats, caption = "Summary Statistics for Telehealth Usage")

```

Analysis of demographic factors revealed that individuals who did not use telehealth in the past year (total = 2,188) outnumbered those who did (total = 1,676). Telehealth users were generally younger (`Age_mean` = 2.93) than non-users (`Age_mean` = 3.03), and they had higher education levels on average (`Education_mean` = 3.49, median = 4) compared to non-users (`Education_mean` = 3.30, median = 4). Income levels also showed a positive association with telehealth usage, as users reported higher average income (`Income_mean` = 5.82) than non-users (`Income_mean` = 5.55). Gender differences were relatively minor, with both male and female participants represented similarly across user groups. However, disparities in telehealth adoption were observed among ethnic groups, with higher usage rates among "White" and "Asian or Pacific Islander" participants compared to "Black" and "Hispanic" groups. These findings suggest that younger age, higher education, and greater income are positively associated with telehealth usage, while disparities among ethnic groups highlight potential barriers to access.

### Social Media Usage Frequency and Health Information Engagement and Telehealth

**please write your results here. Also, mindful in changing the header levels we used for data scripts according to paper components.**

#### RQ2_1. Data Cleaning

```{r}
# Selecting relevant columns for RQ2

data_nl <- hints %>% 
  select(HHID, ReceiveTelehealthCare, SocMed_Visited, 
         SocMed_WatchedVid, SocMed_Interacted, SocMed_SharedPers)

# Renaming each column

data_nl <- data_nl %>% 
  rename(telehealth = ReceiveTelehealthCare, 
         sm_usage_freq = SocMed_Visited, 
         sm_watching = SocMed_WatchedVid, 
         sm_interact =
        SocMed_Interacted, 
        sm_sharing = SocMed_SharedPers)

# Recoding responses

data_nl <- data_nl %>% 
  mutate(telehealth = 
           recode(telehealth, 
                  "No telehealth visits in the past 12 months" = "No", 
                  "Yes, by video" = "Yes", 
                  "Yes, by phone call (voice only with no video)" = "Yes", 
                  "Yes, some by video and some by phone call" = "Yes", 
                  "Missing data (Web partial - Question Never Seen)" = 
                  "Missing", "Missing data (Not Ascertained)" = 
                  "Missing"))

# Creating a dataset for 5 point Likert scale

data_nl_2 <- data_nl %>% 
  mutate(sm_usage_freq = 
           recode(sm_usage_freq, 
                  "Never" = 1, 
                  "Less than once a month" = 2, 
                  "A few times a month" = 3, 
                  "At least once a week" = 4, 
                  "Almost every day" = 5, 
                  "Missing data (Not Ascertained)" = NA_real_, 
                  "Missing data (Web partial - Question Never Seen)" = NA_real_, 
                  "Multiple responses selected in error" = NA_real_), 
         sm_watching = recode(sm_watching, 
                     "Missing data (Not Ascertained)" = NA_real_, 
                     "Missing data (Web partial - Question Never Seen)" = NA_real_, 
                     "Never" = 1, 
                     "Less than once a month" = 2, 
                     "A few times a month" = 3, 
                     "At least once a week" = 4, 
                     "Almost every day" = 5), 
         sm_interact = recode(sm_interact, 
                     "Missing data (Not Ascertained)" = NA_real_, 
                     "Missing data (Web partial - Question Never Seen)" = NA_real_, 
                     "Never" = 1, 
                     "Less than once a month" = 2, 
                     "A few times a month" = 3, 
                     "At least once a week" = 4, 
                     "Almost every day" = 5), 
         sm_sharing = recode(sm_sharing, 
                    "Missing data (Not Ascertained)" = NA_real_, 
                    "Missing data (Web partial - Question Never Seen)" = NA_real_, 
                    "Multiple responses selected in error" = NA_real_, 
                    "Never" = 1, 
                    "Less than once a month" = 2, 
                    "A few times a month" = 3, 
                    "At least once a week" = 4, 
                    "Almost every day" = 5)) %>% 
  filter(!is.na(sm_usage_freq)) %>% 
  filter(!is.na(sm_watching)) %>% 
  filter(!is.na(sm_interact)) %>% 
  filter(!is.na(sm_sharing)) %>% 
  filter(telehealth !="Missing")
```

#### RQ2_2. Grouping by Health Information Engagement Levels on Social Media: pivot_longer()

```{r}
# Reshape health-information-engagement-on-social-media categories into one longer engagement_type column

data_long <- data_nl_2 %>% 
  pivot_longer(cols = c(sm_watching, sm_interact, sm_sharing), 
               names_to = "engagement_type", values_to = "frequency")

# Low, medium, high grouping

data_long <- data_long %>% 
  group_by(HHID) %>% 
  mutate(total_sm_health_eng_lev = sum(frequency))

data_freq_eng_group <- data_long %>% 
  distinct(HHID, telehealth, sm_usage_freq, total_sm_health_eng_lev) %>% 
  mutate(sm_health_eng_lev_group = 
           case_when(
             total_sm_health_eng_lev >= 3 & 
               total_sm_health_eng_lev <= 6 ~ "Low",
             total_sm_health_eng_lev >= 7 & 
               total_sm_health_eng_lev <= 10 ~ "Medium",
             total_sm_health_eng_lev >= 11 & 
               total_sm_health_eng_lev <= 15 ~ "High"))
```

#### RQ2_3. Data Visualization

```{r}
# Social Media Usage Frequency x Telehealth

ggplot(data_freq_eng_group, aes(x = telehealth, fill = as.factor(sm_usage_freq)))+
         geom_bar() +
  scale_fill_discrete(
    labels = c("Never", "Less than once a month", 
               "A few times a month", "At least once a week", 
               "Almost every day")) +
         theme_minimal() +
  labs(
    title = "Social Media Usage Frequency\nAmong Telehealth Users and Non-Users", 
       x = "Telehealth", y = "Count", 
       fill = "Social Media Usage Frequency")

ggplot(data_freq_eng_group, aes(
  x = as.factor(sm_usage_freq), fill = telehealth)) +
  geom_bar(position = "dodge") +
   scale_x_discrete(
    labels = c("Never", "<1/M", 
               "A few/M", "≥1/W", 
               "Every day")) +
        theme_minimal() +
    labs(
      title = "Telehealth Usage by Social Media Usage Frequency",
      x = "Social Media Usage Frequency",
      y = "Count",
      fill = "Telehealth Usage"
    )

# Health Information Engagement Levels on Social Media X Telehealth

ggplot(data_freq_eng_group, aes(x = telehealth, fill = sm_health_eng_lev_group)) +
  geom_bar() +
  theme_minimal() +
  labs(
    title = "Health Information Engagement Levels on Social Media\n Among Telehealth Users and Non-Users", 
    x = "Telehealth", y = "Count", 
    fill = "Health Information Engagement Levels\non Social Media")
  
ggplot(data_freq_eng_group, 
       aes(x = sm_health_eng_lev_group, fill = telehealth)) +
    geom_bar(position = "dodge") +
    theme_minimal() +
    labs(
      title = "Telehealth Usage by Health Information Engagement\nLevels on Social Media", 
      x = "Health Information Engagement Levels\non Social Media",
      y = "Count",
      fill = "Telehealth Usage"
    )
```

#### RQ2_4. Summary Statistics

```{r}
# Social Media Usage Frequency X Telehealth 

summary_table_sm_freq <- data_freq_eng_group %>% 
  mutate(sm_usage_freq = case_when(
      sm_usage_freq == 1 ~ "1 (Never)",
      sm_usage_freq == 2 ~ "2 (Less than once a month)",
      sm_usage_freq == 3 ~ "3 (A few times a month)",
      sm_usage_freq == 4 ~ "4 (At least once a week)",
      sm_usage_freq == 5 ~ "5 (Almost every day)")) %>% 
  group_by(telehealth, sm_usage_freq) %>% 
  summarize(Count = n()) %>% 
  mutate(Percentage = round(Count / sum(Count) * 100, 2)) %>% 
  rename("Telehealth" = telehealth, "SM Usage Freqency" = sm_usage_freq) 

kable(summary_table_sm_freq, caption = "1. Telehealth & Social Media Usage Frequency")

summary_table_sm_freq_2 <- data_freq_eng_group %>% 
  group_by(telehealth, sm_usage_freq) %>% 
  summarize(Count = n()) %>% 
  pivot_wider(names_from = sm_usage_freq, values_from = Count) %>% 
  rename("Telehealth" = telehealth)

kable(summary_table_sm_freq_2, caption = "2. Telehealth & Social Media Usage Frequency")

# Mean, Median, SD for Social Media Usage Frequency X Telehealth

summary_table_sm_freq_stats <- data_freq_eng_group %>% 
  group_by(telehealth) %>% 
  summarize(SM_Usage_Freq_Mean = round(mean(sm_usage_freq), 2), 
            Median = median(sm_usage_freq), 
            SD = round(sd(sm_usage_freq), 2), 
            Count = n()) %>% 
  rename("Telehealth" = telehealth, "Mean" = SM_Usage_Freq_Mean)

kable(summary_table_sm_freq_stats, caption = "Mean, Median, SD for Social Media Usage Frequency by Telehealth")

# Health Information Engagement Levels on Social Media X Telehealth

summary_table_eng_lev_group <- data_freq_eng_group %>% 
  group_by(telehealth, sm_health_eng_lev_group) %>% 
  summarize(Count = n()) %>% 
  mutate(Percentage = round(Count / sum(Count) * 100, 2)) %>% 
  rename(
    "Telehealth" = telehealth, 
    "HI SM Eng" = sm_health_eng_lev_group) 

kable(summary_table_eng_lev_group, caption = "1. Telehealth & HI SM Engagement Level")

summary_table_eng_lev_group_2 <- data_freq_eng_group %>% 
  group_by(telehealth, sm_health_eng_lev_group) %>% 
  summarize(Count = n()) %>% 
  pivot_wider(names_from = sm_health_eng_lev_group, values_from = Count) %>% 
  rename("Telehealth" = telehealth)

kable(summary_table_eng_lev_group_2, caption = "2. Telehealth & HI SM Engagement Level")

# Mean, Median, SD for Health Information Engagement Levels on Social Media X Telehealth

summary_table_eng_lev_stats <- data_long %>% 
  group_by(telehealth) %>% 
  summarize(SM_Health_Info_Eng_Mean = round(mean(total_sm_health_eng_lev), 2), 
            Median = median(total_sm_health_eng_lev), 
            SD = round(sd(total_sm_health_eng_lev), 2)) %>% 
  rename("Telehealth" = telehealth, "Mean" = SM_Health_Info_Eng_Mean) 

kable(summary_table_eng_lev_stats, caption = "Mean, Median, SD for HI SM Engagement Level by Telehealth")
```

### Cancer Risk Perception and Telehealth

Analysis revealed that regardless of cancer risk perception, there were more individuals who had not used telehealth in the past year (total = 2188) compared to those who had used it (total = 1676). Additionally, as cancer risk perception increased, the number of people who had used telehealth decreased (low = 636, medium = 551, high = 489).

```{r}
#| echo: false

# data cleaning

# Selecting relevant columns for RQ3
cancer <- hints %>% 
  select(HHID, ReceiveTelehealthCare, InterestedCaScreening, FreqWorryCancer, ChanceGetCancer2, EverythingCauseCancer, PreventNotPossible)

# Renaming each column
cancer <- cancer %>% 
  rename(telehealth = ReceiveTelehealthCare, 
         screening = InterestedCaScreening, 
         worry = FreqWorryCancer, 
         chance = ChanceGetCancer2, 
         cause = EverythingCauseCancer, 
         prevent = PreventNotPossible)

# Recoding responses
cancer <- cancer %>% 
  mutate(telehealth = fct_collapse(telehealth,
                  "Yes" = c("Yes, by video", "Yes, by phone call (voice only with no video)", "Yes, some by video and some by phone call"), 
                  "No" = "No telehealth visits in the past 12 months", 
                  "Missing" = c("Missing data (Web partial - Question Never Seen)", "Missing data (Not Ascertained)")))

# Creating a dataset for 5 point Likert scale
cancer_2 <- cancer %>% 
  mutate(screening = 
           recode(screening, 
                  "Missing data (Not Ascertained)" = NA_real_, 
                  "Missing data (Web partial - Question Never Seen)" = NA_real_, 
                  "Multiple responses selected in error" = NA_real_, 
                  "Not at all" = 1, 
                  "A little" = 2, 
                  "Somewhat" = 3, 
                  "Very" = 4, 
                  "Not applicable/I am up-to-date with screening tests" = 5), 
         worry = 
           recode(worry, 
                  "Missing data (Not Ascertained)" = NA_real_, 
                  "Missing data (Web partial - Question Never Seen)" = NA_real_,
                  "Multiple responses selected in error" = NA_real_,
                  "Not at all" = 1, 
                  "Slightly" = 2, 
                  "Somewhat" = 3, 
                  "Moderately" = 4, 
                  "Extremely" = 5), 
         chance = 
           recode(chance, 
                  "Missing data (Not Ascertained)" = NA_real_, 
                  "Missing data (Web partial - Question Never Seen)" = NA_real_,
                  "Multiple responses selected in error" = NA_real_,
                  "I already had cancer" = NA_real_,
                  "I don't know" = NA_real_,
                  "Very unlikely" = 1, 
                  "Unlikely" = 2, 
                  "Neither likely nor unlikely" = 3, 
                  "Likely" = 4, 
                  "Very likely" = 5), 
         cause = 
           recode(cause, 
                  "Missing data (Not Ascertained)" = NA_real_, 
                  "Missing data (Web partial - Question Never Seen)" = NA_real_,
                  "Multiple responses selected in error" = NA_real_,
                  "Strongly disagree" = 1, 
                  "Somewhat disagree" = 2, 
                  "Somewhat agree" = 4, 
                  "Strongly agree" = 5), 
         prevent = 
           recode(prevent, 
                  "Missing data (Not Ascertained)" = NA_real_, 
                  "Missing data (Web partial - Question Never Seen)" = NA_real_, 
                  "Multiple responses selected in error" = NA_real_, 
                  "Strongly disagree" = 1, 
                  "Somewhat disagree" = 2, 
                  "Somewhat agree" = 4, 
                  "Strongly agree" = 5)) %>% 
  filter(!is.na(screening)) %>% 
  filter(!is.na(worry)) %>% 
  filter(!is.na(chance)) %>% 
  filter(!is.na(cause)) %>%
  filter(!is.na(prevent)) %>%
  filter(telehealth !="Missing")
```

```{r}
#| echo: false

# grouping by cancer risk perception levels

# Reshape cancer risk perceptions into one longer column
cancer_long <- cancer_2 %>% 
  pivot_longer(
    cols = c(screening, worry, chance, cause, prevent), 
    names_to = "riskperceptiontype", 
    values_to = "riskperceptiondegree"
  )

# Low, medium, high grouping
cancer_long <- cancer_long %>% 
  group_by(HHID) %>% 
  mutate(riskperceptionlevel = mean(riskperceptiondegree, na.rm = TRUE))

# Keeping only one risk perception degree value per household id
cancer_risk_group <- cancer_long %>% 
  distinct(HHID, telehealth, riskperceptionlevel) %>% 
  mutate(risk_group = 
           case_when(
             riskperceptionlevel <= 2.6 ~ "Low", 
             riskperceptionlevel <= 3.2 ~ "Medium", 
             riskperceptionlevel <= 5 ~ "High"))
```

```{r}
#| echo: false

#Data Visualization

# Cancer Risk Levels X Telehealth
cancer_risk_group %>%
  mutate(risk_group = factor(risk_group, levels = c("Low", "Medium", "High"))) %>%
  group_by(risk_group, telehealth) %>%
  summarize(count = n()) %>%
  mutate(percentage = round(count / sum(count) * 100, 2)) %>%
  ggplot(aes(x = risk_group, fill = telehealth, y = count)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste(percentage, "%", sep = "")), 
            position = position_dodge(0.9), vjust = -0.5) +
  theme_minimal() +
  labs(title = "Figure 3.1. Telehealth Usage by Cancer Risk Perception Levels", 
       x = "Cancer Risk Perception Level", 
       y = "Number of Responses", 
       fill = "Telehealth Usage")
```

```{r}
#| echo: false

#summary statistics
# Cancer Risk Levels X Telehealth
summary_table_cancer_risk_group <- cancer_risk_group %>% 
  group_by(telehealth, risk_group) %>% 
  summarize(count = n()) %>% 
  mutate(percentage = round(count/sum(count)*100, 2),
         risk_group = factor(risk_group, levels = c("Low", "Medium", "High")),
         telehealth = factor(telehealth, levels = c("Yes", "No"))) %>%
  arrange(telehealth, risk_group)
kable(summary_table_cancer_risk_group, caption = "Table 3.1. Telehealth Usage by Cancer Risk Perception Levels: Distribution and Proportions")
```

However, individuals who had used telehealth had a higher average cancer risk perception (m = 2.90, sd = 1.29) compared to those who had not used telehealth (m = 2.79, sd = 1.30).

```{r}
#| echo: false

# Mean, Median, SD for SME X Telehealth
summary_table_cancer_risk_stats <- cancer_long %>% 
  group_by(telehealth) %>% 
  summarize(riskperception_mean = round(mean(riskperceptiondegree), 2), 
            riskperception_median = median(riskperceptiondegree), 
            riskperception_sd = round(sd(riskperceptiondegree), 2))
kable(summary_table_cancer_risk_stats, caption = "Table 3.2. Summary Statistics of Cancer Risk Perception by Telehealth Usage")
```

## Discussion

**Demographic Factors and Telehealth Usage**

The findings on demographic factors and telehealth usage revealed that younger individuals, those with higher education levels, and those with greater income are more likely to use telehealth services. However, the lower usage rates among "Black" and "Hispanic" groups highlight disparities that could stem from systemic barriers, including limited access to technology, lower levels of digital literacy, or cultural preferences for in-person care. While telehealth adoption is growing, the demographic disparities suggest that targeted interventions are needed to bridge these gaps.

For example, increasing outreach and education about telehealth benefits in underserved communities, combined with ensuring affordable and reliable access to digital tools, could help enhance adoption [@scott2015]. Additionally, healthcare policies should aim to align telehealth services with the specific needs of different demographic groups [@gallegos-rejas2023]. For individuals with lower income or education levels, providing user-friendly platforms and personalized guidance might improve engagement [@dixit2022]. Meanwhile, culturally sensitive approaches tailored to diverse ethnic groups may foster trust and encourage broader acceptance of telehealth [@willis2022].

**(discussion for RQ2)**

The findings for association between cancer risk perception level and telehealth usage revealed that regardless of the cancer risk perception level, a greater number of individuals had not utilized telehealth compared to those who had. It suggests that telehealth adoption still face some challenges such as accessibility, awareness, or user preferences for in-person care. Especially, considering another result that showed telehealth usage decreased as cancer risk perception levels increased, it can be interpreted that individuals with higher cancer risk perceptions are more likely to seek in-person consultation for perceived severe health concerns. Still, on average, telehealth users reported slightly higher cancer risk perceptions than non-users. This implies that individuals with heightened health concerns are more inclined to use telehealth, possibly as a convenient tool to address their health-related issues. In summary, these results suggest that although telehealth has not yet been fully widespread, future healthcare service policies may utilize telehealth as a complementary tool for face-to-face services. For people with low cancer risk perception, providing healthcare services that actively use telehealth may increase accessibility and efficiency. On the other hand, for people with high cancer concerns, a strategy that provides face-to-face services as the main method while using telehealth as a supplement may be effective. This approach may contribute to devising a plan that can meet the needs of individuals with different levels of risk perception while maximizing the potential of telehealth.

\newpage

# References
