---
title: "Final_Project"
author: "Songyi Ahn, Nakyung Lee, Nishat Parvez"
format:
  html:
   highlight-style: tango
   code-fold: true
   toc: true
   toc-depth: 5
   toc-location: left
execute:
  echo: true
  eval: true
  warning: false
  message: false
editor: visual
---

# Common Work

## Loading Packages

```{r}
#| echo: false
#| include: false

library(tidyverse)
library(here)
library(rio)
library(knitr)
```

## Loading and Renaming the dataset

```{r}
# Load the file
file_path <- here("data", "hints6_public.rda")
load(file_path)

# Change the file name to 'hints'
hints <- public
```

# RQ1. What is the distribution of demographic/socio-economic categories, including age, gender, ethnicity, education, and income level, among individuals who do and do not use Telehealth services?

## RQ1_1. Data Cleaning

```{r}
#this is for Nishat's work 
```

# RQ2. How do social media usage frequency and health information engagement levels on social media (low, medium, high) differ between individuals who use and do not use Telehealth services?

## RQ2_1. Data Cleaning

```{r}
# Selecting relevant columns for RQ2

data_nl <- hints %>% 
  select(HHID, ReceiveTelehealthCare, SocMed_Visited, 
         SocMed_WatchedVid, SocMed_Interacted, SocMed_SharedPers)

# Renaming each column

data_nl <- data_nl %>% 
  rename(telehealth = ReceiveTelehealthCare, 
         sm_usage_freq = SocMed_Visited, 
         sm_watching = SocMed_WatchedVid, 
         sm_interact =
        SocMed_Interacted, 
        sm_sharing = SocMed_SharedPers)

# Recoding responses

data_nl <- data_nl %>% 
  mutate(telehealth = 
           recode(telehealth, 
                  "No telehealth visits in the past 12 months" = "No", 
                  "Yes, by video" = "Yes", 
                  "Yes, by phone call (voice only with no video)" = "Yes", 
                  "Yes, some by video and some by phone call" = "Yes", 
                  "Missing data (Web partial - Question Never Seen)" = 
                  "Missing", "Missing data (Not Ascertained)" = 
                  "Missing"))

# Creating a dataset for 5 point Likert scale

data_nl_2 <- data_nl %>% 
  mutate(sm_usage_freq = 
           recode(sm_usage_freq, 
                  "Never" = 1, 
                  "Less than once a month" = 2, 
                  "A few times a month" = 3, 
                  "At least once a week" = 4, 
                  "Almost every day" = 5, 
                  "Missing data (Not Ascertained)" = NA_real_, 
                  "Missing data (Web partial - Question Never Seen)" = NA_real_, 
                  "Multiple responses selected in error" = NA_real_), 
         sm_watching = recode(sm_watching, 
                     "Missing data (Not Ascertained)" = NA_real_, 
                     "Missing data (Web partial - Question Never Seen)" = NA_real_, 
                     "Never" = 1, 
                     "Less than once a month" = 2, 
                     "A few times a month" = 3, 
                     "At least once a week" = 4, 
                     "Almost every day" = 5), 
         sm_interact = recode(sm_interact, 
                     "Missing data (Not Ascertained)" = NA_real_, 
                     "Missing data (Web partial - Question Never Seen)" = NA_real_, 
                     "Never" = 1, 
                     "Less than once a month" = 2, 
                     "A few times a month" = 3, 
                     "At least once a week" = 4, 
                     "Almost every day" = 5), 
         sm_sharing = recode(sm_sharing, 
                    "Missing data (Not Ascertained)" = NA_real_, 
                    "Missing data (Web partial - Question Never Seen)" = NA_real_, 
                    "Multiple responses selected in error" = NA_real_, 
                    "Never" = 1, 
                    "Less than once a month" = 2, 
                    "A few times a month" = 3, 
                    "At least once a week" = 4, 
                    "Almost every day" = 5)) %>% 
  filter(!is.na(sm_usage_freq)) %>% 
  filter(!is.na(sm_watching)) %>% 
  filter(!is.na(sm_interact)) %>% 
  filter(!is.na(sm_sharing)) %>% 
  filter(telehealth !="Missing")
```

## RQ2_2. Grouping by Health Information Engagement Levels on Social Media: pivot_longer()

```{r}
# Reshape health-information-engagement-on-social-media categories into one longer engagement_type column

data_long <- data_nl_2 %>% 
  pivot_longer(cols = c(sm_watching, sm_interact, sm_sharing), 
               names_to = "engagement_type", values_to = "frequency")

# Low, medium, high grouping

data_long <- data_long %>% 
  group_by(HHID) %>% 
  mutate(total_sm_health_eng_lev = sum(frequency))

data_freq_eng_group <- data_long %>% 
  distinct(HHID, telehealth, sm_usage_freq, total_sm_health_eng_lev) %>% 
  mutate(sm_health_eng_lev_group = 
           case_when(
             total_sm_health_eng_lev >= 3 & 
               total_sm_health_eng_lev <= 6 ~ "Low",
             total_sm_health_eng_lev >= 7 & 
               total_sm_health_eng_lev <= 10 ~ "Medium",
             total_sm_health_eng_lev >= 11 & 
               total_sm_health_eng_lev <= 15 ~ "High"))
```

## RQ2_3. Data Visualization

```{r}
# Social Media Usage Frequency x Telehealth

ggplot(data_freq_eng_group, aes(x = telehealth, fill = as.factor(sm_usage_freq)))+
         geom_bar() +
  scale_fill_discrete(
    labels = c("Never", "Less than once a month", 
               "A few times a month", "At least once a week", 
               "Almost every day")) +
         theme_minimal() +
  labs(
    title = "Social Media Usage Frequency\nAmong Telehealth Users and Non-Users", 
       x = "Telehealth", y = "Count", 
       fill = "Social Media Usage Frequency")

ggplot(data_freq_eng_group, aes(
  x = as.factor(sm_usage_freq), fill = telehealth)) +
  geom_bar(position = "dodge") +
   scale_x_discrete(
    labels = c("Never", "<1/M", 
               "A few/M", "≥1/W", 
               "Every day")) +
        theme_minimal() +
    labs(
      title = "Telehealth Usage by Social Media Usage Frequency",
      x = "Social Media Usage Frequency",
      y = "Count",
      fill = "Telehealth Usage"
    )

# Health Information Engagement Levels on Social Media X Telehealth

ggplot(data_freq_eng_group, aes(x = telehealth, fill = sm_health_eng_lev_group)) +
  geom_bar() +
  theme_minimal() +
  labs(
    title = "Health Information Engagement Levels on Social Media\n Among Telehealth Users and Non-Users", 
    x = "Telehealth", y = "Count", 
    fill = "Health Information Engagement Levels\non Social Media")

data_freq_eng_group$sm_health_eng_lev_group <- factor(
  data_freq_eng_group$sm_health_eng_lev_group, 
  levels = c("Low", "Medium", "High"))
  
ggplot(data_freq_eng_group, 
       aes(x = sm_health_eng_lev_group, fill = telehealth)) +
    geom_bar(position = "dodge") +
    theme_minimal() +
    labs(
      title = "Telehealth Usage by Health Information Engagement\nLevels on Social Media", 
      x = "Health Information Engagement Levels\non Social Media",
      y = "Count",
      fill = "Telehealth Usage"
    )
```

## RQ2_4. Summary Statistics

```{r}
# Social Media Usage Frequency X Telehealth 

summary_table_sm_freq <- data_freq_eng_group %>% 
  mutate(sm_usage_freq = case_when(
      sm_usage_freq == 1 ~ "1 (Never)",
      sm_usage_freq == 2 ~ "2 (Less than once a month)",
      sm_usage_freq == 3 ~ "3 (A few times a month)",
      sm_usage_freq == 4 ~ "4 (At least once a week)",
      sm_usage_freq == 5 ~ "5 (Almost every day)")) %>% 
  group_by(telehealth, sm_usage_freq) %>% 
  summarize(Count = n()) %>% 
  mutate(Percentage = round(Count / sum(Count) * 100, 2)) %>% 
  rename(Telehealth = telehealth, SM_Usage_Freqency = sm_usage_freq) 

kable(summary_table_sm_freq)

# Mean, Median, SD for Social Media Usage Frequency X Telehealth

summary_table_sm_freq_stats <- data_freq_eng_group %>% 
  group_by(telehealth) %>% 
  summarize(SM_Usage_Freq_Mean = round(mean(sm_usage_freq), 2), 
            Median = median(sm_usage_freq), 
            SD = round(sd(sm_usage_freq), 2), 
            Count = n()) %>% 
  rename(Telehealth = telehealth)

kable(summary_table_sm_freq_stats)

# Health Information Engagement Levels on Social Media X Telehealth

summary_table_eng_lev_group <- data_freq_eng_group %>% 
  group_by(telehealth, sm_health_eng_lev_group) %>% 
  summarize(Count = n()) %>% 
  mutate(Percentage = round(Count / sum(Count) * 100, 2)) %>% 
  rename(
    Telehealth = telehealth, 
    SM_Health_Info_Eng = sm_health_eng_lev_group) 

kable(summary_table_eng_lev_group)

# Mean, Median, SD for Health Information Engagement Levels on Social Media X Telehealth

summary_table_eng_lev_stats <- data_long %>% 
  group_by(telehealth) %>% 
  summarize(SM_Health_Info_Eng_Mean = round(mean(total_sm_health_eng_lev), 2), 
            Median = median(total_sm_health_eng_lev), 
            SD = round(sd(total_sm_health_eng_lev), 2)) %>% 
  rename(Telehealth = telehealth) 

kable(summary_table_eng_lev_stats)
```

# RQ3. How does the distribution of cancer-related risk perception groups (low, medium, high) compare between individuals who use and do not use Telehealth services?

## RQ3_1. Data Cleaning

```{r}
# Selecting relevant columns for RQ3
cancer <- hints %>% 
  select(HHID, ReceiveTelehealthCare, InterestedCaScreening, FreqWorryCancer, ChanceGetCancer2, EverythingCauseCancer, PreventNotPossible)

# Renaming each column
cancer <- cancer %>% 
  rename(telehealth = ReceiveTelehealthCare, 
         screening = InterestedCaScreening, 
         worry = FreqWorryCancer, 
         chance = ChanceGetCancer2, 
         cause = EverythingCauseCancer, 
         prevent = PreventNotPossible)

# Recoding responses
cancer <- cancer %>% 
  mutate(telehealth = 
           recode(telehealth, 
                  "No telehealth visits in the past 12 months" = "No", 
                  "Yes, by video" = "Yes", 
                  "Yes, by phone call (voice only with no video)" = "Yes", 
                  "Yes, some by video and some by phone call" = "Yes", 
                  "Missing data (Web partial - Question Never Seen)" = "Missing", 
                  "Missing data (Not Ascertained)" = "Missing"))

# Creating a dataset for 5 point Likert scale
cancer_2 <- cancer %>% 
  mutate(screening = 
           recode(screening, 
                  "Missing data (Not Ascertained)" = NA_real_, 
                  "Missing data (Web partial - Question Never Seen)" = NA_real_, 
                  "Multiple responses selected in error" = NA_real_, 
                  "Not at all" = 1, 
                  "A little" = 2, 
                  "Somewhat" = 3, 
                  "Very" = 4, 
                  "Not applicable/I am up-to-date with screening tests" = 5), 
         worry = 
           recode(worry, 
                  "Missing data (Not Ascertained)" = NA_real_, 
                  "Missing data (Web partial - Question Never Seen)" = NA_real_,
                  "Multiple responses selected in error" = NA_real_,
                  "Not at all" = 1, 
                  "Slightly" = 2, 
                  "Somewhat" = 3, 
                  "Moderately" = 4, 
                  "Extremely" = 5), 
         chance = 
           recode(chance, 
                  "Missing data (Not Ascertained)" = NA_real_, 
                  "Missing data (Web partial - Question Never Seen)" = NA_real_,
                  "Multiple responses selected in error" = NA_real_,
                  "I already had cancer" = NA_real_,
                  "I don't know" = NA_real_,
                  "Very unlikely" = 1, 
                  "Unlikely" = 2, 
                  "Neither likely nor unlikely" = 3, 
                  "Likely" = 4, 
                  "Very likely" = 5), 
         cause = 
           recode(cause, 
                  "Missing data (Not Ascertained)" = NA_real_, 
                  "Missing data (Web partial - Question Never Seen)" = NA_real_,
                  "Multiple responses selected in error" = NA_real_,
                  "Strongly disagree" = 1, 
                  "Somewhat disagree" = 2, 
                  "Somewhat agree" = 4, 
                  "Strongly agree" = 5), 
         prevent = 
           recode(prevent, 
                  "Missing data (Not Ascertained)" = NA_real_, 
                  "Missing data (Web partial - Question Never Seen)" = NA_real_, 
                  "Multiple responses selected in error" = NA_real_, 
                  "Strongly disagree" = 1, 
                  "Somewhat disagree" = 2, 
                  "Somewhat agree" = 4, 
                  "Strongly agree" = 5)) %>% 
  filter(!is.na(screening)) %>% 
  filter(!is.na(worry)) %>% 
  filter(!is.na(chance)) %>% 
  filter(!is.na(cause)) %>%
  filter(!is.na(prevent)) %>%
  filter(telehealth !="Missing")
```

## RQ3_2. Grouping by Cancer Risk Perception Groups

```{r}
# Reshape cancer risk perceptions into one longer column
cancer_long <- cancer_2 %>% 
  pivot_longer(
    cols = c(screening, worry, chance, cause, prevent), 
    names_to = "riskperceptiontype", 
    values_to = "riskperceptiondegree"
  )

# Low, medium, high grouping
cancer_long <- cancer_long %>% 
  group_by(HHID) %>% 
  mutate(riskperceptionlevel = mean(riskperceptiondegree, na.rm = TRUE))

# Keeping only one risk perception degree value per household id
cancer_risk_group <- cancer_long %>% 
  distinct(HHID, telehealth, riskperceptionlevel) %>% 
  mutate(risk_group = 
           case_when(
             riskperceptionlevel <= 1.67 ~ "Low", 
             riskperceptionlevel <= 3.33 ~ "Medium", 
             riskperceptionlevel <= 5 ~ "High"))
```

## RQ3_3. Data Visualization

```{r}
# Cancer Risk Levels X Telehealth
cancer_risk_group %>%
  mutate(risk_group = factor(risk_group, levels = c("Low", "Medium", "High"))) %>%
  ggplot(aes(x = risk_group, fill = telehealth)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(title = "Telehealth Usage by Cancer Risk Perception Levels", 
       x = "Cancer Risk Perception Level", 
       y = "Number of Responses", 
       fill = "Telehealth Usage")
```

## RQ3_4. Summary Statistics

```{r}
# Cancer Risk Levels X Telehealth
summary_table_cancer_risk_group <- cancer_risk_group %>% 
  group_by(telehealth, risk_group) %>% 
  summarize(count = n()) %>% 
  mutate(percentage = round(count/sum(count)*100, 2))
kable(summary_table_cancer_risk_group)

# Mean, Median, SD for SME X Telehealth
summary_table_cancer_risk_stats <- cancer_long %>% 
  group_by(telehealth) %>% 
  summarize(riskperception_mean = round(mean(riskperceptiondegree), 2), 
            riskperception_median = median(riskperceptiondegree), 
            riskperception_sd = round(sd(riskperceptiondegree), 2))
kable(summary_table_cancer_risk_stats)
```
